\chapter{Conclusion} % (4-6 pages)
\label{chap:conclusion}

\section{Challenges}
\noindent
Several challenges were faced during the course of this project, the main ones being:

Firstly, it was challenging to find information on the technical aspects of existing Version Control Systems, especially the earlier generations of Version Control Systems. The lack of information on these technical aspects made it difficult to understand how the systems were implemented and what underlying data structures and algorithms were used.
\smallskip

Secondly, it was challenging to identify which data structures and algorithms would provide meaningful points of comparison. The vast number of data structures and algorithms made it challenging to select a suitable set for evaluation. The selected data structures and algorithms had to be representative of those used in real-world Version Control Systems and provide a basis for meaningful comparison.
\smallskip

Thirdly, I had to learn the Go programming language's built-in benchmarking package, which they were initially unfamiliar with. The learning curve for this package was steep, and it took time to understand how to use it effectively for the purposes of this research project.
\smallskip

Finally, it was challenging to design automated benchmarks using mock data that were still representative of real-world use. The mock data had to be generated in a way that was realistic and representative of the data used in real-world Version Control Systems. This required careful consideration of the different types of data used in Version Control Systems and how they could be represented in the mock data.

\section{Future Work}
\noindent
One possible area for future work is implementing a fully functional Version Control System to obtain the most realistic data for benchmarking. The fully functional Version Control System would allow for more realistic testing and benchmarking of the data structures and algorithms and provide insight into their performance under real-world usage scenarios.
Other areas for future work are evaluating other data structures and algorithms not considered in this research project and developing more advanced benchmarking techniques than those used in this project.